Internship Summary: Mid-Term Review

### 2020-07-28 RapportStage/Rapport.tex
fix issue with phi

The eigenvectors associated to the eigenenergies \eqref{eq:E_2x2} are
\begin{equation}\label{eq:phi_2x2}
\begin{split}
	\phi_{\pm}(\lambda)
	& = 
	\begin{pmatrix}
		(\epsilon_1 - \epsilon_2 \pm \sqrt{(\epsilon_1 - \epsilon_2)^2 + 4\lambda^2})/2\lambda 
		\\ 
		1
	\end{pmatrix}
	\\
	& =
	\begin{pmatrix}
		(\epsilon_1-\epsilon_2)/2\lambda \pm \sqrt{2\lambda_\text{EP}} \sqrt{\lambda - \lambda_\text{EP}}/\lambda 
		\\ 
		1
	\end{pmatrix},
\end{split}
\end{equation}
and, for $\lambda=\lambda_\text{EP}$, they become
\begin{align}
	\phi_{\pm}\qty(i\,\frac{\epsilon_1 - \epsilon_2}{2}) & = \begin{pmatrix} -i \\ 1\end{pmatrix},
	&
	\phi_{\pm}\qty(-i\,\frac{\epsilon_1 - \epsilon_2}{2}) & = \begin{pmatrix} i \\ 1\end{pmatrix}.
\end{align}
which are clearly self-orthogonal, i.e., their norm is equal to zero. 
%Using Eq.~\eqref{eq:E_EP}, Eq.~\eqref{eq:phi_2x2} can be recast as
%\begin{equation}\label{eq:phi_EP}
%\phi_{\pm}=
%\end{equation}
As branch point (square-root) singularities, EPs inherit their topology from the multi-valued function $\sqrt{\lambda - \lambda_\text{EP}}$. 
Then, if the eigenvectors are properly normalized, they behave as $(\lambda - \lambda_\text{EP})^{-1/4}$, which gives the following pattern when looping around one EP:
\begin{align}
	\phi_{\pm}(2\pi) & = \phi_{\mp}(0),
	&
	\phi_{\pm}(4\pi) & = -\phi_{\pm}(0), 
	\\
	\phi_{\pm}(6\pi) & = -\phi_{\mp}(0),
	&
	\phi_{\pm}(8\pi) & = \phi_{\pm}(0).
\end{align}
In plain words, four loops around the EP are necessary to recover the initial state. 
We can also see that looping the other way around leads to a different pattern.

@@

In the Hartree-Fock (HF) approximation, the many-electron wave function is approximated as a single Slater determinant $\Psi^{\text{HF}}(\vb{x}_1,\ldots,\vb{x}_n)$ [where $\vb{x} = (\sigma,\vb{r})$ is a composite vector gathering spin and spatial coordinates] defined as an antisymmetric combination of $n$ (real-valued) one-electron spin-orbitals $\phi_p(\vb{x})$, which are, by definition, eigenfunctions of the one-electron Fock operator 
\begin{equation}\label{eq:FockOp}
	f(\vb{x}) \phi_p(\vb{x}) = [ h(\vb{x}) + v^\text{HF}(\vb{x}) ] \phi_p(\vb{x}) = \epsilon_p \phi_p(\vb{x}),
\end{equation}
where 
\begin{equation}
	h(\vb{x}) = -\frac{\grad^2}{2} + \sum_{A}^{N} \frac{Z_A}{\abs{\vb{r}-\vb{R}_A}}
\end{equation}
is the core Hamiltonian and 
\begin{equation}
	v^\text{HF}(\vb{x}) = \sum_i \qty[ J_i(\vb{x}) - K_i(\vb{x}) ]
\end{equation}
is the HF mean-field potential with 
\begin{gather}
	\label{eq:CoulOp}
	J_i(\vb{x})\phi_p(\vb{x})=\qty[\int\dd\vb{x}'\phi_i(\vb{x}')\frac{1}{\abs{\vb{r} - \vb{r}'}}\phi_i(\vb{x}') ] \phi_p(\vb{x})
	\\
	\label{eq:ExcOp}
	K_i(\vb{x})\phi_p(\vb{x})=\qty[\int\dd\vb{x}'\phi_i(\vb{x}')\frac{1}{\abs{\vb{r} - \vb{r}'}}\phi_p(\vb{x}') ] \phi_i(\vb{x})
\end{gather}
being the Coulomb and exchange operators (respectively) in the spin-orbital basis.
From hereon, $i$ and $j$ are occupied orbitals, $a$ and $b$ are unoccupied (or virtual) orbitals, while $p$, $q$, $r$, and $s$ indicate arbitrary orbitals.
Rather than solving Eq.~\eqref{eq:SchrEq}, HF theory uses the variational principle to find an approximation of $\Psi$ as a single Slater determinant. Hence a Slater determinant is not an eigenfunction of the exact Hamiltonian $\hH$. However, it is, by definition, an eigenfunction of the so-called (approximated) HF many-electron Hamiltonian defined as the sum of the one-electron Fock operators
\begin{equation}\label{eq:HFHamiltonian}
	\hH^{\text{HF}} = \sum_{i}^{n} f(\vb{x}_i).
\end{equation}
Note that the lowest eigenvalue of the Hamiltonian \eqref{eq:HFHamiltonian} is not the HF energy but the sum of the eigenvalues associated with the occupied eigenvalues (see below).

@@

\begin{equation}\label{eq:MPHamiltonian}
    \hH(\lambda) = 
    \sum_{i}^{n} \qty[ 
    -\frac{\grad_i^2}{2} 
    - \sum_{A}^{N} \frac{Z_A}{\abs{\vb{r}_i-\vb{R}_A}} 
    + (1-\lambda) v^{\text{HF}}(\vb{x}_i) 
    + \lambda\sum_{i<j}^{n}\frac{1}{\abs{\vb{r}_i-\vb{r}_j}} 
    ].
\end{equation}
As mentioned earlier, in perturbation theory, the energy is a power series of $\lambda$ and the physical energy is obtained at $\lambda = 1$. 
The MP$n$ energy is defined as 
\begin{equation}
	E_{\text{MP}n}= \sum_{k=0}^n E^{(k)},
\end{equation}
where $E^{(k)}$ is the $k$th-order correction.
The MP0 energy overestimates the energy by double counting the electron-electron interaction and is equal to the sum of the occupied orbital energies, i.e.,
\begin{equation}
	E_{\text{MP0}} = \sum_i^n \epsilon_i.
\end{equation}
The MP1 corrects this and is then equal to the HF energy, i.e.,
\begin{equation}
	E_{\text{MP1}} = E_\text{HF} = \sum_i^n \epsilon_i - \frac{1}{2} \sum_{ij}^n \mel{ij}{}{ij},
\end{equation}
with $\mel{pq}{}{rs} = \braket{pq}{rs} - \braket{pq}{sr}$, and where  
\begin{equation}
	\braket{pq}{rs} = \iint \dd\vb{x}_1\dd\vb{x}_2\phi_p(\vb{x}_1)\phi_q(\vb{x}_2)\frac{1}{\abs{\vb{r}_1 - \vb{r}_2}}\phi_r(\vb{x}_1)\phi_s(\vb{x}_2)
\end{equation}
is a two-electron integral in the spin-orbital basis.
MP2 starts recovering correlation energy and the MP2 energy, which reads
\begin{equation}\label{eq:EMP2}
	E_{\text{MP2}} = \frac{1}{4} \sum_{ij} \sum_{ab} \frac{\abs{\mel{ij}{}{ab}}^2}{\epsilon_i + \epsilon_j - \epsilon_a - \epsilon_b},
\end{equation}
is then lower than the HF energy \cite{SzaboBook}.

>>>>

The eigenvectors associated to the eigenenergies \eqref{eq:E_2x2} are
\begin{equation}\label{eq:phi_2x2}
\begin{split}
	\phi_{\pm}
	& = 
	\begin{pmatrix}
		(\epsilon_1 - \epsilon_2 \pm \sqrt{(\epsilon_1 - \epsilon_2)^2 + 4\lambda^2})/2\lambda 
		\\ 
		1
	\end{pmatrix}
	\\
	& =
	\begin{pmatrix}
		(\epsilon_1-\epsilon_2)/2\lambda \pm \sqrt{2\lambda_\text{EP}} \sqrt{\lambda - \lambda_\text{EP}}/\lambda 
		\\ 
		1
	\end{pmatrix},
\end{split}
\end{equation}
and, for $\lambda=\lambda_\text{EP}$, they become
\begin{align}
	\phi_{\pm} & = \begin{pmatrix} -i \\ 1\end{pmatrix},
	&
	\phi_{\pm} & = \begin{pmatrix} i \\ 1\end{pmatrix},
\end{align}
which are clearly self-orthogonal, i.e., their norm is equal to zero. 
%Using Eq.~\eqref{eq:E_EP}, Eq.~\eqref{eq:phi_2x2} can be recast as
%\begin{equation}\label{eq:phi_EP}
%\phi_{\pm}=
%\end{equation}
As branch point (square-root) singularities, EPs inherit their topology from the multi-valued function $\sqrt{\lambda - \lambda_\text{EP}}$. 
Then, if the eigenvectors are properly normalized, they behave as $(\lambda - \lambda_\text{EP})^{-1/4}$, which gives the following pattern when looping around one EP:
\begin{align}
	\phi_{\pm}(2\pi) & = \phi_{\mp}(0),
	&
	\phi_{\pm}(4\pi) & = -\phi_{\pm}(0), 
	&
	\phi_{\pm}(6\pi) & = -\phi_{\mp}(0),
	&
	\phi_{\pm}(8\pi) & = \phi_{\pm}(0).
\end{align}
In plain words, four loops around the EP are necessary to recover the initial state. 
We can also see that looping the other way around leads to a different pattern.

@@

In the Hartree-Fock (HF) approximation, the many-electron wave function is approximated as a single Slater determinant $\Psi^{\text{HF}}(\vb{x}_1,\ldots,\vb{x}_n)$ [where $\vb{x} = (\sigma,\vb{r})$ is a composite vector gathering spin and spatial coordinates] defined as an antisymmetric combination of $n$ (real-valued) one-electron spin-orbitals $\phi_p(\vb{x})$, which are, by definition, eigenfunctions of the one-electron Fock operator 
\begin{equation}\label{eq:FockOp}
	f(\vb{x}) \phi_p(\vb{x}) = [ h(\vb{x}) + v^\text{HF}(\vb{x}) ] \phi_p(\vb{x}) = \epsilon_p \phi_p(\vb{x}),
\end{equation}
where 
\begin{equation}
	h(\vb{x}) = -\frac{\grad^2}{2} + \sum_{A}^{N} \frac{Z_A}{\abs{\vb{r}-\vb{R}_A}}
\end{equation}
is the core Hamiltonian and 
\begin{equation}
	v^\text{HF}(\vb{x}) = \sum_i \qty[ J_i(\vb{x}) - K_i(\vb{x}) ]
\end{equation}
is the HF mean-field potential with 
\begin{subequations}
\begin{gather}
	\label{eq:CoulOp}
	J_i(\vb{x})\phi_p(\vb{x})=\qty[\int\dd\vb{x}'\phi_i(\vb{x}')\frac{1}{\abs{\vb{r} - \vb{r}'}}\phi_i(\vb{x}') ] \phi_p(\vb{x})
	\\
	\label{eq:ExcOp}
	K_i(\vb{x})\phi_p(\vb{x})=\qty[\int\dd\vb{x}'\phi_i(\vb{x}')\frac{1}{\abs{\vb{r} - \vb{r}'}}\phi_p(\vb{x}') ] \phi_i(\vb{x})
\end{gather}
\end{subequations}
being the Coulomb and exchange operators (respectively) in the spin-orbital basis.
From hereon, $i$ and $j$ are occupied orbitals, $a$ and $b$ are unoccupied (or virtual) orbitals, while $p$, $q$, $r$, and $s$ indicate arbitrary orbitals.
Rather than solving Eq.~\eqref{eq:SchrEq}, HF theory uses the variational principle to find an approximation of $\Psi$ as a single Slater determinant. Hence a Slater determinant is not an eigenfunction of the exact Hamiltonian $\hH$. However, it is, by definition, an eigenfunction of the so-called (approximated) HF many-electron Hamiltonian defined as the sum of the one-electron Fock operators
\begin{equation}\label{eq:HFHamiltonian}
	\hH^{\text{HF}} = \sum_{i}^{n} f(\vb{x}_i).
\end{equation}
Note that the lowest eigenvalue of the Hamiltonian \eqref{eq:HFHamiltonian} is not the HF energy but the sum of the eigenvalues associated with the occupied eigenvalues (see below).

@@

\begin{equation}\label{eq:MPHamiltonian}
    \hH(\lambda) = 
    \sum_{i}^{n} \qty[ 
    -\frac{\grad_i^2}{2} 
    - \sum_{A}^{N} \frac{Z_A}{\abs{\vb{r}_i-\vb{R}_A}} 
    + (1-\lambda) v^{\text{HF}}(\vb{x}_i) 
    + \lambda\sum_{i<j}^{n}\frac{1}{\abs{\vb{r}_i-\vb{r}_j}} 
    ].
\end{equation}
As mentioned earlier, in perturbation theory, the energy is a power series of $\lambda$ and the physical energy is obtained at $\lambda = 1$. 
The MP$n$ energy is defined as 
\begin{equation}
	E_{\text{MP}n}= \sum_{k=0}^n E^{(k)},
\end{equation}
where $E^{(k)}$ is the $k$th-order correction.
\titou{$n$ is the number of electrons and here you use it as something else...}
The MP0 energy overestimates the energy by double counting the electron-electron interaction and is equal to the sum of the occupied orbital energies, i.e.,
\begin{equation}
	E_{\text{MP0}} = \sum_i^n \epsilon_i.
\end{equation}
The MP1 corrects this and is then equal to the HF energy, i.e.,
\begin{equation}
	E_{\text{MP1}} = E_\text{HF} = \sum_i^n \epsilon_i - \frac{1}{2} \sum_{ij}^n \mel{ij}{}{ij},
\end{equation}
with $\mel{pq}{}{rs} = \braket{pq}{rs} - \braket{pq}{sr}$, and where  
\begin{equation}
	\braket{pq}{rs} = \iint \dd\vb{x}_1\dd\vb{x}_2\phi_p(\vb{x}_1)\phi_q(\vb{x}_2)\frac{1}{\abs{\vb{r}_1 - \vb{r}_2}}\phi_r(\vb{x}_1)\phi_s(\vb{x}_2)
\end{equation}
is a two-electron integral in the spin-orbital basis.
MP2 starts recovering correlation energy and the MP2 energy, which reads
\begin{equation}\label{eq:EMP2}
	E_{\text{MP2}} = \frac{1}{4} \sum_{ij} \sum_{ab} \frac{\abs{\mel{ij}{}{ab}}^2}{\epsilon_i + \epsilon_j - \epsilon_a - \epsilon_b},
\end{equation}
is then lower than the HF energy \cite{SzaboBook}.
